---
title: "Exploring Emails"
author: "Jonathan Roberts"
date: "7 April 2017"
output: 
    html_notebook: default
    html_document: default
---

#Get up and running
##Global RMarkdown Parameters
Are set here

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

##Install the R markdown packages and library

```{r message = FALSE, warning = TRUE}
libraries <- c("rmarkdown")

#UNCOMMENT TO INSTALL PACKAGES
#lapply(libraries, install.packages)
lapply(libraries, library, character.only = TRUE)
```


## Working Directory: 
Set up where you are working and what is in the directory and what objects you have
 
```{r}

# SET THE WORKING DIRECTORY
setwd("/Users/jonathanroberts/Documents/R/Email_Analysis")
dir()
ls()

```

#Assignment: Part 1 & 2


#Playing Around with Email Data

Long term aim might be to do:
topic modelling and see how the changes over time
LSA/clustering and see how that changes over time
Maybe doing that sort of thing when passing a set of users over to it

Another area to explore might be linking to other people's data and having some network diagrams perhaps.

Ultimately i'd like to get something which might be a jonbot and suggest responses to emails for me...



##Write in the data and have a look at it

```{r}

raw_data <- read.csv("../SensitiveData/2017_sent_emails.CSV", colClasses = "character")
str(raw_data)
summary(raw_data)
class(raw_data)
colnames(raw_data)
raw_data[,6]

#outcome[, 11] <- as.numeric(outcome[, 11])
## You may get a warning about NAs being introduced; that is okay

```


```{r}

best <- function(state,outcome){
  
  ##Read in the data
  outcomes <- read.csv("hospital_data/outcome-of-care-measures.csv", colClasses = "character")

  ##Set the columns where the data is - and the master list of conditions
  conditions<-data.frame( c("heart attack","heart failure","pneumonia"),c(11,17,23))

  #find out if the call is valid
  state_test<-is.na(match(state,outcomes$State)) ## returns TRUE for an ERROR
  condition_test<- is.na(match(outcome,conditions[,1])) ## returns TRUE for an ERROR
  valid_call<-condition_test | state_test ## returns TRUE for an ERROR

  ##TODO - WRITE THE STOP LINE TO EXIT THE FUNCTION
  if (valid_call == TRUE) {
      stop("Data was wrong") 
      }

  
  
  #Make my data numeric
  
  #data<-conditions[,2]
  ####outcomes[,data]<-as.numeric(outcomes[,data]) ##TODO - why doesn't this work??
  outcomes[,11]<-as.numeric(outcomes[,11])
  outcomes[,17]<-as.numeric(outcomes[,17])  
  outcomes[,23]<-as.numeric(outcomes[,23])
  
  # SPLIT BY STATE: THIS IS QUITE RESOURCE INTENSIVE - MIGHT BE BEST NOT TO 'FILTER' ON STATE NAME.
  outcome_by_state<-split(outcomes, outcomes$State)
  #PUL OUT THE SELECTED STATE DAATA AS A DATA FRAME
  state_df<-outcome_by_state[[state]]
  
  #READY TO 'SORT' - ORDER SEEMS TO BE THE BEST...
  cond<-conditions[match(outcome,conditions[,1]),2]  
  order_by_MR<-state_df[order(state_df[,cond],state_df[,2]),]
  order_by_MR[1,2]
  
}

  best("TX","heart failure")

```

#Sandbox: Part 1 &2

##Write in the data

```{r}

outcome <- read.csv("hospital_data/outcome-of-care-measures.csv", colClasses = "character")
head(outcome)
str(outcome)
summary(outcome)
class(outcome)
colnames(outcome)
outcome[,11]
ncol(outcome)
nrow(outcome)
outcome[, 11] <- as.numeric(outcome[, 11])
## You may get a warning about NAs being introduced; that is okay
hist(outcome[, 11])
head(outcome)

```

## Do the main bit of ordering and finding number 1
```{r}
#Where is the data?
heart_attack<-outcome[,11]

outcome[,17]<-as.numeric(outcome[,17])
heart_failure<-outcome[,17]
hist(outcome[, 17])

outcome[,23]<-as.numeric(outcome[,23])
pneumonia<-outcome[,23]
hist(outcome[, 23])


# SPLIT BY STATE: THIS IS QUITE RESOURCE INTENSIVE - MIGHT BE BEST NOT TO SPLIT AND 'FILTER' ON STATE NAME.
outcome_by_state<-split(outcome, outcome$State)
class(outcome_by_state)
str(outcome_by_state)
#PUL OUT THE SELECTED STATE DAATA AS A DATA FRAME
class(outcome_by_state$AK)
ak_df<-outcome_by_state$AK
ak_df

#READY TO 'SORT' - ORDER SEEMS TO BE THE BEST...

order_by_MR<-ak_df[order(ak_df[,11],ak_df[,2]),]
order_by_MR[1,]

```

##Do the error trapping
```{r}
#find out if we are one of the states

state<-"AK"
condition<-"heart attack"
conditions<-data.frame( c("heart attack","heart failure","pneumonia"),c(11,17,23))

state_test<-is.na(match(state,outcomes$State)) ## returns TRUE for an ERROR
condition_test<- is.na(match(condition,conditions[,1])) ## returns TRUE for an ERROR

valid_call<-condition_test | state_test ## returns TRUE for an ERROR

if (valid_call == True) {print("You entering something incorrectly")}



```



